%
% Copyright 1992-1998 by Werner Almesberger.
% All rights reserved.
%
% See file COPYING for details.
%

\documentstyle[fullpage]{article}

\parindent=0pt
\parskip=4pt

\hyphenation{hexa-decimal}

\def\key#1{$[$#1$]$}
\def\LILO{LILO}

\def\SetFigFont#1#2#3{\tt}


\begin{document}

\title{\LILO \\
  \Large
  Generic boot loader for Linux \\
  Version 21 \\ ~\\
  \bf Technical overview}
\author{Werner Almesberger \\
  {\tt Werner.Almesberger@epfl.ch}}
\date{December 4, 1998}

\maketitle
{
  \parskip=-4pt
  \setcounter{tocdepth}{1}
  \tableofcontents
}

~\\
This document describes internals of \LILO\ and related parts of its
environment (kernel, etc.). It is not necessary to read or understand
this document in order to install or use \LILO. A general introduction
and installation instructions can be found in the user's guide.

{\bf This document has only been partially updated and does not entirely
reflect the current ('98) status of Linux or of \LILO\ (version 21).}


\section{Load sequence}
\label{load}

The boot sector is loaded by the ROM-BIOS at address 0x07C00. It moves
itself to address 0x96A00, sets up the stack (growing downwards from
0x96A00 to 0x96800), loads the secondary boot loader at address
0x96C00 and transfers control to it. It displays an ``L'' after moving
itself and an ``I'' before starting the secondary boot loader. If a read
error occurs when loading the secondary boot loader, a two-digit hex code
is displayed after the ``L''. This results in an endless stream of error
codes if the problem is permanent. Displaying these error codes is disabled
if the build-time option {\tt NO1STDIAG} is set.

The secondary boot loader loads the descriptor table at 0x98800 and the
sector containing the default command line at 0x98C00. If the default command
line is enabled, its magic number is invalidated and the sector is written
back to disk. This potentially dangerous operation can be disabled by defining
{\tt LCF\_READONLY} when passing {\tt second.S} through {\sf cpp}. Next, the
secondary boot loader checks
for user input. If either the default is used or if the user has specified
an alternate image, the options sector is loaded at 0x98C00 and the parameter
line is constructed at 0x99000. If the resulting line contains the option
\verb"lock", the command line as entered by the user (it is saved before
the final line is constructed) is written to the disk as the new default
command line. Also, if a fallback command line is set, it is copied to the
default command line sector.

If the user has supplied an initial RAM disk image, this file is loaded
below the end of physical memory or 16 MB, whichever is lower. The start
address is lowered to the next page boundary so that the memory area
occupied by the initial RAM disk can later be easily returned to the
system's free memory pool. The 16 MB limit exists because the BIOS
functions used to transfer data in memory are only specified for an 24 bit
address space.

Next, the floppy boot sector of that image is
loaded at 0x90000\footnote{The floppy boot sector is only used as a source of
setup information.},
the setup part is loaded at 0x90200 and
the kernel part is loaded at 0x10000, or, if the kernel has been compiled
for being loaded ``high'' (i.e. with {\tt make bzImage}), it is loaded
at 0x100000 instead. During the load operations, the sectors
of the map file are loaded at 0x98600.

If the loaded image is a kernel image, control is transferred to
its setup code.
If a different operating system is booted, things are a bit more difficult:
the chain loader is loaded at 0x90200 and the boot sector of the other OS
is loaded at 0x90400. The chain loader moves the partition table (loaded at
0x903BE as part of the chain loader) to 0x00600 and the boot sector to
0x07C00. After that, it passes control to the boot sector.

Chain loaders that allow booting from a second drive (either floppy or hard
disk) also install a small function to intercept BIOS calls and to swap
the drive numbers at the top of available memory.

The secondary boot loader displays an ``L'' after being started and an ``O''
after loading the descriptor table and the default command line. Before
loading the descriptor table, it checks, whether it has been loaded at the
correct location and displays a question mark if it hasn't. If the
descriptor table has an incorrect checksum, a minus sign is displayed.

$$
\begin{tabular}{l|c|l}
  \cline{2-2}
  \tt 0x00000 & & 1982 bytes \\
  \cline{2-2}
  \tt 0x007BE & Partition table & 64 bytes \\
  \cline{2-2}
  \tt 0x007FE & & 29 kB \\
  \cline{2-2}
  \tt 0x07C00 & Boot load area & 512 bytes \\
  \cline{2-2}
  \tt 0x07E00 & & 32.5 kB \\
  \cline{2-2}
  \tt 0x10000 & & 448 kB \\
  & & \\
  & Kernel & \\
  & & \\
  & & \\
  \cline{2-2}
  \tt 0x90000 & Floppy boot sector & 512 bytes \\
  \cline{2-2}
  \tt 0x90200 & Setup (kernel) & 39.5 kB (2 kB used) \\
  \cline{2-2}
  \tt 0x9A000 & Primary boot loader & 512 bytes \\
  \cline{2-2}
  \tt 0x9A200 & Stack & 3.5 kB \\
  \cline{2-2}
  \tt 0x9B000 & Secondary boot loader & 8 kB (3.5 kB used) \\
  \cline{2-2}
  \tt 0x9D000 & Map load area & 512 bytes \\
  \cline{2-2}
  \tt 0x9D200 & Descriptor table & 1 kB \\
  \cline{2-2}
  \tt 0x9D600 & Default command line, etc. & 512 bytes \\
  \cline{2-2}
  \tt 0x9D800 & Keyboard translation table & 512 bytes \\
  \cline{2-2}
  \tt 0x9DA00 & Parameter line construction area & 1 kB \\
  \cline{2-2}
  \tt 0x9DC00 & & 7.5 kB \\
  & & \\
  \cline{2-2}
  & Drive swapper & 1 kB \\
  \cline{2-2}
  \multicolumn{3}{l}{\tt 0xA0000} \\
\end{tabular}
$$

The area 0x90020-0x90023 is overlaid by a command-line descriptor while
the secondary boot loader is running.



\section{File references}

This section describes the references among files involved in the boot
procedures.

$$
  \input bootloader
$$

The boot sector contains the primary boot loader, the address of the default
command line sector, the address of both
descriptor table sectors and the addresses of the sectors of the secondary
boot loader. The generic boot sector is copied from {\tt boot.b}.

The primary boot loader can store up to eight sector addresses of the
secondary boot loader.

$$
  \input map
$$

The map file consists of so-called sections and of special data sectors.
Each section
spans an integral number of disk sectors and contains addresses of sectors
of other files.

There are three exceptions: 1. If a ``hole'' is being
covered or if the floppy boot sector of an unstripped kernel has been omitted,
the address of the zero sector is used. This sector is part of the
map file. 2. When booting a different operating system, the first sector is
the merged chain loader that has been written to the map file before that
section. 3. Each map section describing an image is followed by a sector
containing the options line of that image.

The last address slot of each map sector is either unused (if the map section
ends in this sector) or contains the address of the next map sector in the
section.

The ifive sectors at the beginning of the map file are special: the first
sector contains the default command line, the next
two sectors contain the boot image descriptor table and the fourth sector
is filled with zero bytes. This sector is mapped whenever a file contains
a ``hole''. The fifth sector contains the keyboard translation table.

$$
  \input image
$$

A kernel image consists simply of a sequence of sectors being
loaded. The map section also contains a sector with a fallback command line
and a sector with parameter line options. Optionally, a RAM disk image,
specified by a second map section, can be loaded.

$$
  \input other
$$

When booting another operating system, the chain loader ({\tt chain.b}) is
merged with the patched partition table\footnote{If the partition table is
omitted, that area is filled with zero bytes.} and written into the map file.
The map section of this boot image starts after that sector and contains only
the address of a dummy floppy boot sector (the zero sector, but its
contents are irrelevant), the loader
sector and the boot sector of the other operating system. Not that the
map section also contains the fallback sector and a (useless) sector for
options.


\section{Configuration parameters}

The boot sector of each kernel contains a set of configuration parameters
that have to be available at boot time before the kernel can access
file systems. These parameters can be set when the kernel is compiled and
later be changed with programs like {\sf rdev}. \LILO\ can supersede
the parameters (in memory) at boot time by placing the corresponding
items on the parameter line passed to the kernel.

The parameters are stored at the following (decimal) offsets:

\begin{description}
  \item[497] the size of the setup code in sectors (512 bytes). Older kernels
    may put a zero at this place.
  \item[498-499] is a flag specifying whether the root file system should be
    mounted read-only (if non-zero) or read-write (if zero).
  \item[500-501] the size of the kernel, counted in paragraphs (16 bytes).
  \item[502-503] this parameter is currently unused.
  \item[504-505] the size of the RAM disk in kilobytes. No RAM disk is
    created if this parameter is set to zero.
  \item[506-507] the text mode the VGA is set to.
    \begin{description}
      \item[0xFFFD] the user is asked to specify the VGA mode at boot time.
      \item[0xFFFE] uses 80x50 (``extended'') mode.
      \item[0xFFFF] uses 80x25 (``normal'') mode.
    \end{description}
    Any other value selects the corresponding mode as displayed in the
    interactive VGA mode selection menu. This is the only option that is set
    by LILO by patching the boot sector instead of passing it on the parameter
    line.
  \item[508] the minor number of the device that should be mounted as root.
  \item[509] the major number of the device that should be mounted as root.
\end{description}


\section{Parameter line interface}

The kernel supports processing of parameters that are
provided by the boot loader. The parameter string is a NUL-terminated
ASCII string that contains space-separated words or
{\tt {\it variable\/}={\it value\/}} pairs. A description of how they are
interpreted can be found in the section of the user's guide labeled
``The boot prompt''.

The following descriptor has to be set up to pass a parameter string to
the kernel:

\begin{description}
  \item[0x90020] the magic number 0xA33F.
  \item[0x90022] the offset of the first byte of the parameter line relative
    to 0x90000.
\end{description}

The boot loader composes the parameter line from the command line, from
the options sector and from some internally generated prefixes (typically
\verb"auto" and \verb"BOOT_IMAGE="), as follows:

$$
  \input parameter
$$

Example: \\
Command line: \verb"vmlinuz root=802"\\
Options sector: \verb"root=801 ro"

yields \verb"BOOT_IMAGE=vmlinuz root=801 ro root=802"

Because parameter line options can typically be overridden, the first
\verb"root" option is ignored by the kernel.


\section{External interface}

\LILO\ is able to receive its command line from a program that is booted
before it. This externally provided command line is only used if the
user does not use the normal mechanism to invoke the boot prompt.

The following register contents are expected:

\begin{description}
  \item[\bf DL] contains the value 0xFE.
  \item[\bf ES:SI] points to the string ``LILO''. The string must be in
    upper case and no terminating character is needed. The string must not
    cross segment boundaries, i.e. {\bf SI} must be below 0xFFFD.
  \item[\bf ES:BX] points to a NUL-terminated string that is used as the
    command line. This string has a maximum length of 78 characters (not
    including the terminating NUL) and must not cross segment boundaries.
\end{description}

There are two values of the externally provided command line that have a
special meaning:

\begin{itemize}
  \item an empty string ({\bf ES:BX} points to a NUL byte) is interpreted
    as a request to enter the boot prompt and to accept keyboard input.
  \item a string that consists only of blanks is interpreted as a request
    to boot the default boot image.
\end{itemize}

\LILO\ can also obtain the default command line from the map file. It
is only used if no externally provided command line is available.


\section{Default command line in map file}

The first sector of the map file is reserved for a default command line.
Unless the user invokes the boot prompt by pressing a shift key or unless an
externally provided command line is present, the command line in the map
file is interpreted as if it had been typed on the keyboard.

The first two bytes of the first sector of the map file have to contain
the magic number {\tt DC\_MAGIC} (0xF4F2) in little-endian byte order.
They are followed by a NUL-terminated string with a maximum length of
510 bytes, including the NUL. Note that the boot loader limits command
lines to 78 characters after removing duplicate spaces.

The command line is disabled by either clobbering the magic number or
by using an empty string (i.e. only a NUL byte) as the command line.

\end{document}
